Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-23 cav2017/cav2017.tex
quantiZation

The approach is proven sound by accounting for errors due to the digitalisation of signals that are manipulated by the controller block.  
The proposed synthesis approaches and a developed tool are evaluated considering 
a set of intricate physical plant models taken from the digital control literature.
\keywords{
State-space dynamical models of physical systems; 
Digital controllers; 
Analogue-to-digital converters; 
Time sampling; 
Quantisation; 
Fixed-point arithmetic; 
CEGIS; 
safety requirements. 
}
\end{abstract}

@@

Linear Time Invariant (LTI) control models represent a broad class of dynamical systems
%modern embedded software, performing highly non-trivial control tasks,
with significant impact in numerous application areas such as
the life sciences, robotics, and engineering~\cite{astrom1997computer,Franklin15}.
Control synthesis for LTI is broadly investigated, 
however the use of digital control architectures adds new challenges due to 
%exacerbated by specific artefacts,  
the effects of finite-precision arithmetic, time discretization, and quantisation noise, which are typically introduced 
by Analogue-to-Digital (A/D) and Digital-to-Analogue (D/A) conversion.  
Whilst the research on digital control is well developed~\cite{astrom1997computer}, 
automated and sound control synthesis is a challenging task. 
%Thus, the programming is a key barrier to broad adoption of digital control, 
%and requires considerable expertise.

@@

We consider models of physical plants expressed as ordinary differential equations (ODEs), 
with full state information (namely where we have access to all the model variables):  
%
\begin{align}
\label{eq:ode}
\dot{x}(t) = Ax(t)+ B u(t), \quad x \in \mathbb{R}^{n}, u \in \mathbb{R}^m, A \in \mathbb{R}^{n \times n}, B \in \mathbb{R}^{n \times m}, 
\end{align}
where $t \in \mathbb R_0^+$, 
where $A$ and $B$ are matrices that fully specify the continuous plant, 
and with initial states set as $x(0)$. 
Eq.~\eqref{eq:ode} is soundly discretized in time (as shown in the Appendix) into %\textcolor{red}{What is subscript ``d''?}
%
\begin{align}
\label{eq:plant}
x_{k+1} = A_d x_k+ B_d u_k. 
\end{align} 
where $k \in \mathbb N$, 
and with initial state $x_{0}=x(0)$. $A_d$ and $B_d$ denote the discretized matrices of $A$ and $B$, the plant dynamics. 
Later, we will deal with the issue of variable quantisation, as introduced by the ADC and DAC blocks.

@@

%\addtodo{[Here discuss Jury criterium for DISCRETE-TIME models - a simplified version of the first part of what is now in Section \ref{sec:cof_verification} (with no quantisation noises yet).]}

@@

Thus far we have presented a verification procedure for a digitalised plant in a continuous space (\emph{i.e.} the
range of the quantisers is still $\mathbb{R}$. The next step we must realise is that the controller 
exists in a digital program, hence its domain is $\mathbb{R}\langle I,F\rangle$. We have $\mat{A},\mat{B},\vec{x} \in \mathbb{R}^*$, $\mat{K}_d,\vec{u},\hat{\vec{x}}=\mathcal{F}\langle I,F \rangle(\vec{x}) \in \mathbb{R}\langle I,F\rangle^*$
Where $*$ represents the appropriate dimension for each matrix/vector.
Unfortunately this means that although the closed loop dynamics are in the reals, they suffer from
the fixed word restrictions in their composing elements.
Apart from the quantisation errors $q_1,q_2$ there are rounding errors in the calculation of $u_k$. 
These will introduce a new source of quantisation noise which we may denote $\vec{\nu}_3 \in [-q_3\ \ q_3]$.
The error in $\vec{u}$ corresponds of the multiplication of $\mat{K}_d\hat{\vec{x}}$, hence $q_3=\overline{\delta}_u=(|\mat{K}|_1+1)c_m$

>>>>

The approach is proven sound by accounting for errors due to the digitalisation of signals that are manipulated by the controller block.  
The proposed synthesis approaches and a developed tool are evaluated considering 
a set of intricate physical plant models taken from the digital control literature.
\keywords{
State-space dynamical models of physical systems; 
Digital controllers; 
Analogue-to-digital converters; 
Time sampling; 
quantization; 
Fixed-point arithmetic; 
CEGIS; 
safety requirements. 
}
\end{abstract}

@@

Linear Time Invariant (LTI) control models represent a broad class of dynamical systems
%modern embedded software, performing highly non-trivial control tasks,
with significant impact in numerous application areas such as
the life sciences, robotics, and engineering~\cite{astrom1997computer,Franklin15}.
Control synthesis for LTI is broadly investigated, 
however the use of digital control architectures adds new challenges due to 
%exacerbated by specific artefacts,  
the effects of finite-precision arithmetic, time discretization, and quantization noise, which are typically introduced 
by Analogue-to-Digital (A/D) and Digital-to-Analogue (D/A) conversion.  
Whilst the research on digital control is well developed~\cite{astrom1997computer}, 
automated and sound control synthesis is a challenging task. 
%Thus, the programming is a key barrier to broad adoption of digital control, 
%and requires considerable expertise.

@@

We consider models of physical plants expressed as ordinary differential equations (ODEs), 
with full state information (namely where we have access to all the model variables):  
%
\begin{align}
\label{eq:ode}
\dot{x}(t) = Ax(t)+ B u(t), \quad x \in \mathbb{R}^{n}, u \in \mathbb{R}^m, A \in \mathbb{R}^{n \times n}, B \in \mathbb{R}^{n \times m}, 
\end{align}
where $t \in \mathbb R_0^+$, 
where $A$ and $B$ are matrices that fully specify the continuous plant, 
and with initial states set as $x(0)$. 
Eq.~\eqref{eq:ode} is soundly discretized in time (as shown in the Appendix) into %\textcolor{red}{What is subscript ``d''?}
%
\begin{align}
\label{eq:plant}
x_{k+1} = A_d x_k+ B_d u_k. 
\end{align} 
where $k \in \mathbb N$, 
and with initial state $x_{0}=x(0)$. $A_d$ and $B_d$ denote the discretized matrices of $A$ and $B$, the plant dynamics. 
Later, we will deal with the issue of variable quantization, as introduced by the ADC and DAC blocks.

@@

%\addtodo{[Here discuss Jury criterium for DISCRETE-TIME models - a simplified version of the first part of what is now in Section \ref{sec:cof_verification} (with no quantization noises yet).]}

@@

Thus far we have presented a verification procedure for a digitalised plant in a continuous space (\emph{i.e.} the
range of the quantizers is still $\mathbb{R}$. The next step we must realise is that the controller 
exists in a digital program, hence its domain is $\mathbb{R}\langle I,F\rangle$. We have $\mat{A},\mat{B},\vec{x} \in \mathbb{R}^*$, $\mat{K}_d,\vec{u},\hat{\vec{x}}=\mathcal{F}\langle I,F \rangle(\vec{x}) \in \mathbb{R}\langle I,F\rangle^*$
Where $*$ represents the appropriate dimension for each matrix/vector.
Unfortunately this means that although the closed loop dynamics are in the reals, they suffer from
the fixed word restrictions in their composing elements.
Apart from the quantization errors $q_1,q_2$ there are rounding errors in the calculation of $u_k$. 
These will introduce a new source of quantization noise which we may denote $\vec{\nu}_3 \in [-q_3\ \ q_3]$.
The error in $\vec{u}$ corresponds of the multiplication of $\mat{K}_d\hat{\vec{x}}$, hence $q_3=\overline{\delta}_u=(|\mat{K}|_1+1)c_m$
